---
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>",
    fig.path = "man/figures/README-",
    out.width = "100%",
    message = FALSE,
    warning = FALSE
)

desc = read.dcf("DESCRIPTION")
desc = setNames(as.list(desc), colnames(desc))
```

# `r desc$Package`

<!-- [![CRAN status](https://www.r-pkg.org/badges/version/tabstats)](https://CRAN.R-project.org/package=tabstats) -->
<!-- [![R-CMD-check](https://github.com/joshuamarie/tabstats/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/joshuamarie/tabstats/actions/workflows/R-CMD-check.yaml) -->
<!-- [![Codecov test coverage](https://codecov.io/gh/joshuamarie/tabstats/graph/badge.svg)](https://app.codecov.io/gh/joshuamarie/tabstats) -->

## Package overview

Title: ***`r desc$Title`***

***`r desc$Package`*** is a lightweight package in action that assists you to print out the data you want to display as a table, with more enchantments that customizes table with styles and alignment. This is especially useful in displaying statistical results in R console/command line as a table.

> The package is currently **under construction** and not yet available on CRAN.

## Installation

You can install the development version of `tabstats` from GitHub with:

```r
# install.packages("devtools")
devtools::install_github("joshuamarie/tabstats")
```

## Example

Here’s a basic example that demonstrates how to use `tabstats` to format and style a table:

```{r example}
library(tabstats)

mtcars |> 
    tibble::rownames_to_column() |> 
    table_default(
        justify_cols = list("1" = "right"),
        style_colnames = list(
            "mpg" = function(ctx) cli::col_red(cli::style_bold(ctx$formatted_value)),
            "cyl" = "blue_italic"
        ),
        style_columns = list(
            "mpg" = function(ctx) {
                val <- as.numeric(ctx$value)
                if (val > 20) cli::col_green(ctx$formatted_value)
                else if (val > 15) cli::col_yellow(ctx$formatted_value)
                else cli::col_red(ctx$formatted_value)
            }
        ),
        digits = 1,
        center_table = TRUE,
        n_space = 2
    )
```

Changing the default `border_char` value:

```{r}
mtcars |> 
    tibble::rownames_to_column() |> 
    table_default(
        justify_cols = list("1" = "right"),
        style_colnames = list(
            "mpg" = function(ctx) cli::col_red(cli::style_bold(ctx$formatted_value)),
            "cyl" = "blue_italic"
        ),
        style_columns = list(
            "mpg" = function(ctx) {
                val <- as.numeric(ctx$value)
                if (val > 20) cli::col_green(ctx$formatted_value)
                else if (val > 15) cli::col_yellow(ctx$formatted_value)
                else cli::col_red(ctx$formatted_value)
            }
        ),
        digits = 1,
        border_char = "=",
        center_table = TRUE,
        n_space = 2
    )
```

This package allows users to apply advanced styling and alignment to tables, such as color coding, column alignment, and borders, making it easier to view large datasets directly in the command line.

