---
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>",
    fig.path = "man/figures/README-",
    out.width = "100%",
    message = FALSE,
    warning = FALSE
)

desc = read.dcf("DESCRIPTION")
desc = setNames(as.list(desc), colnames(desc))

library(tabstats)
```

# `r desc$Package`

<!-- [![CRAN status](https://www.r-pkg.org/badges/version/tabstats)](https://CRAN.R-project.org/package=tabstats) -->
[![R-CMD-check](https://github.com/joshuamarie/tabstats/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/joshuamarie/tabstats/actions/workflows/R-CMD-check.yaml)
<!-- [![Codecov test coverage](https://codecov.io/gh/joshuamarie/tabstats/graph/badge.svg)](https://app.codecov.io/gh/joshuamarie/tabstats) -->

## Package overview

Title: ***`r desc$Title`***

***`r desc$Package`*** is a lightweight package in action that assists you to print out the data you want to display as a table, with more enchantments that customizes table with styles and alignment. This is especially useful in displaying statistical results in R console/command line as a table.

> The package is currently **under construction** and not yet available on CRAN.

## Installation

You can install the development version of `tabstats` from GitHub with:

``` r
# install.packages("pak")
pak::pak("joshuamarie/tabstats")
```

## Example

Here’s a basic example that demonstrates how to use `tabstats` to format and style a table:

```{r example}
table_default(head(mtcars, 5))
```

This package allows end users to apply advanced styling and alignment to tables, such as color coding, column alignment, and borders. 

## License

`r desc$License`

## Code of Conduct

Please note that the `tabstats` project is released with a [Contributor Code of Conduct](https://contributor-covenant.org/version/2/1/CODE_OF_CONDUCT.html). By contributing to this project, you agree to abide by its terms.


